### 2014-01-15 figures/fig2/text.tex
ms text figures

\documentclass{article}
\usepackage{graphics}
\usepackage{nopageno}
%\usepackage{txfonts}
\usepackage[usenames]{color}
%\usepackage[T2A]{fontenc}
\usepackage[utf8]{inputenc}
\usepackage{amsmath}
\usepackage{amssymb} %maths
\begin{document}
\begin{center}
\input{fig.tex}
\end{center}
The ratio of the~$F_{Hadoop}/F_{HPC}$ as a function of a reciprocal dataset size in Gigabases.  Calculations were carried out for~$p=56$ and $p=8$ cores for Hadoop and HPC correspondingly.
The points are fit in a line using least-squares method, what makes  possible to make some careful predictions for the case of the {\it infinite} dataset size.
\end{document}