### 2014-01-15 figures/fig8/text.tex

\documentclass{article}
\usepackage{graphics}
\usepackage{nopageno}
\usepackage{txfonts}
\usepackage[usenames]{color}
\usepackage[T2A]{fontenc}
\usepackage[utf8]{inputenc}
%\usepackage{amsmath}

\begin{document}
\begin{center}
\input{fig.tex}
\end{center}
The effeciency ratios\ref{eq:efficiency} $E_{p}/E_{56}$ for different  Hadoop on the Cloud cluster sizes~112  and~224 cores.
The data for Shared~TM and SSH~TM deployment scenarios are shown.
The efficinency ratios are shown for different pipeline stages: reads preprocessing~(1-3), alingment~(4), and SNP calling~(5).
The values above~1 shows that the cluster of current size is more efficient than the $p=56$ cores cluster, and vice versa. 
\end{document}