### 2014-01-15 figures/fig9/text.tex

\documentclass{article}
\usepackage{graphics}
\usepackage{nopageno}
\usepackage{txfonts}
\usepackage[usenames]{color}
\usepackage[T2A]{fontenc}
\usepackage[utf8]{inputenc}
%\usepackage{amsmath}

\begin{document}
\begin{center}
\input{fig.tex}
\end{center}
Ratios of the mapping time~$T_{mapping}$ to the communication costs~$T_{comm}$ for HPC2 and Hadoop clusters for Dataset IV as a function of reciprocal cluster size $1/N$. Two HPC scenarios are shown as ``HPC SLURM'' and ``HPC random'', which correspond to standard SLURM behaviour and a modified one, where nodes are being allocated from random racks.
%One  can see the differences in the Hadoop and HPC communication patterns: for Hadoop the ratio is almost constant, while for HPC cluster there are two defined regions.
%The first one resembles to that of Hadoop, and corresponds to the case then the communication in the cluster is not a bottleneck. 
%The second one shows the regime corresponding to the case of saturation of the network, resulting in traffic jams: data distibution time~$T_{comm}$ stops to scale 
%linearly to the reciprocal cluster size, and becomes more of less a constant, while the mapping time continues to satisfy~$T_{mapping}\propto 1/N$.
Linear fit  done with the least-squares method.
\end{document}