### 2014-04-16 bioinformatics/supplement.tex
Added supplement and updated for alexey comments

\documentclass[11pt]{article}

\usepackage[colorlinks]{hyperref}

\begin{document}

\title{Supplemental information}
\author{}
\date{}
\maketitle

\subsection*{Datasets}

The datasets used in the paper are publicly available at:

I: \url{http://1001genomes.org/data/software/shoremap/shoremap\_2.0\\/data/reads/Schneeberger.2009/Schneeberger.2009.single\_end.gz}

II: \url{http://1001genomes.org/data/software/shoremap/shoremap\_2.0/data/reads/Galvao.2012/Galvao.2012.reads1.fq.gz, http://1001genomes.org/data/software/shoremap/shoremap\_2.0/data/reads/Galvao.2012/Galvao.2012.reads2.fq.gz}

III: \url{ftp://ftp-trace.ncbi.nlm.nih.gov/sra/sra-instant/reads/ByRun/sra/SRR/SRR611/SRR611084//SRR611084.sra, ftp://ftp-trace.ncbi.nlm.nih.gov/sra/sra-instant/reads/ByRun/sra/SRR/SRR611/SRR611085//SRR611085.sra}

IV: artificial pair-ended dataset for {\it A.thaliana} created with the {\tt wgsim} program from the Samtools package.

V: \url{http://www.ncbi.nlm.nih.gov/sra/SRX148888}

\subsection*{Reference genomes}
\begin{itemize}
\item TAIR10 for datasets II-IV \url{ftp://ftp.arabidopsis.org/home/tair/Sequences/whole\_chromosomes/*.fas}
\item TAIR8 for dataset I \url{ftp://ftp.arabidopsis.org/home/tair/Genes/TAIR8\_genome\_release/}
\item H.sapiens, NCBI v37 \url{ftp://ftp.ccb.jhu.edu/pub/data/bowtie\_indexes/h\_sapiens\_37\_asm.ebwt.zip}
\end{itemize}

\subsection*{Description of computational facilities}

\begin{enumerate}
%\item  HPC:
%The HPC analysis pipeline was run on a node from the Kalkyl~\cite{kalkyl} cluster, equipped with two quad-core processors Intel Xeon~5520 (clock frequency of 2.26\,GHz; 1\,MB L2 cache, 8\,MB L3 cache), 24\,GB of RAM and an Infiniband node-to-node network connection, and 10Gbasesit/s uplink. The data and reference genomes were read and written to a parallel shared storage system.

\item 
HPC:
Multinode short-read alignment was performed on the Milou cluster (\url{http://www.uppmax.uu.se/the-milou-cluster}), equipped with dual 8-core Intel Xeon E5-2660, (2.2 GHz, 2\,MB L2 cache, 20\,MB L3 cache), 128\,GB of RAM, Infiniband node-to-node network connection, and 10Gbit/s uplink.

\item Storage: 
Gulo (\url{http://www.uppmax.uu.se/gulo}) is a custom built Lustre 2.4 system using 8 HP nodes with MDS600
storage boxes and an additional node for metadata handling. In total, it
provides roughly 1 PB of storage and is accessed with Lustre's own protocol. It
supports data striping over multiple nodes and disk targets and can give a
theoretical single file read performance of up to 80 Gbits per second.

\item Our Hadoop test platform was deployed on a private cloud at UPPMAX using
	the OpenNebula (\url{http://opennebula.org}) cloud management system. Each physical node
	was equipped with two 4-core Intel Xeon 5420 (2.50\,GHz; 12~MB L2 cache),
	16~GB RAM, one 1~TB SATA disk and Gigabit Ethernet. The cluster was set up
	with Cloudera Hadoop Distribution version~2.0.0-cdh4.5.0 (\url{http://www.cloudera.com/content/cloudera/en/why-cloudera/hadoop-and-big-data.html}).
	Note that the physical hardware provided less RAM than desired. Each VM node
	has 7 cores and 14~GB of RAM (2~GB/core), which is less than recommended for
	Hadoop. We would expect to see much better performance with twice as much memory.
\end{enumerate}

\subsection*{Crossbow preprocessing stage as a Bash script}

The following Bash script mimics functionality of the Crossbow's preprocessing stage,
and uses multi-core parallelism. The script reformats the
\texttt{bzip2}-compressed short-reads in FASTQ format into a single text file
where each line contains following tab-separated fields:
\begin{itemize}
\item read header
\item forward read
\item forward read qualities
\item reverse read
\item reverse read qualities
\end{itemize}

\texttt {pbzip2 -dc  \${file1} | paste - - - - -d'\textbackslash t' | cut -f1,2,4 | paste - -d' ' <(pbzip2 -dc \${file2}) 
| paste - - - - -d'\textbackslash t' | cut -f2,4) | pbzip2 -cz > \${fileOut}}

\end{document}